Coherent IR-Hyper-Raman Four Wave Mixing Spectroscopy

### 2024-08-29 paper.tex
some words

\documentclass[aip, jcp, reprint, onecolumn]{revtex4-2}

@@

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares them with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require macroscopic non-centrosymmetry under the electric dipole approximation.\cite{RN133}
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to FWM cascades.\cite{RN515, RN243, Cho2000_Cascade}
A coherent six-wave mixing analogue of hyper-Raman scattering, coherent anti-Stokes hyper-Raman spectroscopy (CAHRS), has been proposed and explored theoretically by several authors; however, it is also highly susceptible to FWM cascades.\cite{Berger1978, Bjarnason1980, Cho1997, Cho1998}
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling). %rpm: not sure what you're trying to say by this
With the other four-wave and six-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just a single vibrational mode and electronic state (``evx'' coupling).\cite{RN445, RN335} 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as two-color experiments (with $\omega_2=\omega_3$);\cite{Cho2001} vibrational sum frequency generation (vSFG) or difference frequency generation (vDFG) setups can be trivially reconfigured to perform HSFG or HDFG measurements, respectively, to investigate ev coupling in bulk systems.

When the hyper-Raman excitation is resonant or near-resonant with electronic states, the brightness of the vibrational feature depends on the vibronic structure of the electronic state.
The electronic spectrum probed in HDFG can inform on processes that control ultrafast electronic relaxation in molecular and biological systems in similar ways to resonance Raman.\cite{Bredenbeck2015, Arsenault2021}
Since the selection rules of hyper-Raman scattering differ from Raman scattering, the hyper-Raman excitation spectra give a unique alternative to Raman scattering to understand vibrational spectra, electronic structure and vibronic coupling in molecular systems. \cite{Olson2018}

@@

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
HSFG will not be discussed, but the application of our treatment to HSFG is straightforward and the selection rules are similar.
We note that HDFG emission can be phase-matched in media with normal dispersion, while HSFG cannot,\cite{RN278} although experimental strategies can be implemented to minimize phase-mismatch in HSFG processes. \cite{Kaufman2024}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or obscure coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66 in]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is vibrationally resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned. \cite{RN287, McDonnell2024}

@@

It is useful to expose relationships between nonlinear output, transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit. \cite{Simpson2004}
Under the electric dipole approximation, the $I^\text{th}$ cartesian component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation) \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(\omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the $IJKL$ element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}\label{eq:nfgamma}
	\chi^{(3)}_{IJKL} = NF \langle \gamma_{ijkl} \rangle.
\end{equation}
Here $N$ is a number density and $\gamma_{ijkl}$ is the third-order polarizability (i.e., second hyperpolarizability).
The brackets indicate an orientational average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.
$F=\prod_j  \frac{1}{3} \left({\left(n(\omega_j)\right)^2 + 1} \right)$ is the Lorentz local field factor, where $n$ is the (frequency dependent) refractive index. \cite{Bedeaux1973}

@@

\subsection{Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the general source of HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to those found in Albrecht's treatment of Raman spectroscopy.\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(Q)) \otimes \ket{b}$ for electronic states $\{|a(Q))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where $Q$ is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on $Q$ for simplicity.

For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments; $\vec{\mu}$ is reserved for transitions on the ground electronic state. \cite{Tang1970}
Following the approach which gave \autoref{sivegamma} and using the time ordering and state labeling of \autoref{fig:sivewmel2}a, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{4D \hbar^3} \sum_{m,n,e,v'} \frac{
		R_{i}^{gv, ev'} 
		R_{j}^{ev',mn} 
		R_{k}^{mn,g0} 
		R_{l}^{g0,gv} 
	}{\Delta_{g0,gv}
		\Delta_{ev', mn}
		\Delta_{mn, g0}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.
It is common in the hyper-Raman community to write $\Delta_{ev', mn} \Delta_{mn, g0} = \Delta_{ev', g0}$.
Using our definition of vibronic states, we can write, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$, where $\vec{M}_{ab} = (a|\vec{R}|b)$.\cite{Ziegler1974}
Expanding $\vec{M}^{ij}$ to $\order{Q}$ about the equilibrium point of the ground state potential surface as
$\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z} Q_z$
yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Warshel1977, Ziegler1988} so that
\begin{equation}\label{gammaabc}
		\gamma_{ijkl} \sim \left(A_{ijk} + B_{ijk} + C_{ijk}\right) \frac{\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}
\end{equation}
The $A$ term contains the static ($\order{Q^0}$) transitions (i.e., Condon approximation), the $B$ term depends upon $\order{Q}$ transitions (Herzberg-Teller contributions), and the $C$ term depends on $\order{Q^2}$ transitions. 
The $C$ term is suppressed in the following discussion as it depends on one and two photon forbidden transitions, making its contribution to $\gamma_{ijkl}$ roughly two orders of magnitude less than $A_{ijk}$. \cite{Ziegler1988, Neddersen1989, Bonang1992}
Note that some reports obtain $A, B$ coefficients through a Herzberg-Teller expansion of the electronic states to expose vibronic couplings via $\partial H / \partial Q$, where $H$ is the electronic Hamiltonian.\cite{HerzbergTeller1933, Petrov1985, Neddersen1989, Baranov1990}
This approach is not used here as the expansion of $\vec{M}^{ij}$ in normal mode coordinates provides sufficient physical insight into hyper-Raman selection rules.

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$ are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev', g0}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | v' \rangle \mel{v'}{Q_z}{0} \frac{1}{\Delta_{ev', g0}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev', g0}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
where $\langle a | b \rangle$ and $\mel{a}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively. 
Duschinsky rotation effects are ignored in this discussion. \cite{Duschinsky1937}
At this point, the expressions are valid for both electronically resonant and non-resonant cases.
For the case of electronic resonance, only a few components of this summation will be relevant ($e,v^\prime$ such that $\Delta_{ev^\prime,g0}$ is small).

To understand the vibronic structure of electronic resonance, we restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite ground state vibrational modes through $\omega_1$; we will assume the IR beam has excited the fundamental of an IR-active mode so that hyper-Raman scattering is between this mode and the ground state. 
By working only in terms of this single normal mode (i.e., removing the sum over $z$ and taking $v=1$), 
the $A$ and $B$ terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
			\begin{split}
				A_{ijk} = \frac{1}{\hbar}\sum_{v'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | v' \rangle
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
				\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{1_{ijk}} &= \frac{1}{\hbar} \sum_{v'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | v' \rangle \mel{v'}{Q}{0} 
				\frac{1}{\Delta_{ev', g0}}\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{v'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{v'} 
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',g0}}
			\end{split}
		\end{equation}
	\end{subequations}
where $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} $ is the electronic two photon transition moment between $a$ and $b$.\cite{McClain1977}

\autoref{ABterms_DR} shows important selection rules for the vibronic transitions.
The $A$ term is allowed whenever an electronic transition has a one- and two- photon transition moment.
For centrosymmetric molecules, an electronic transition cannot be both one- and two- photon allowed by parity, so the $A$ term will vanish.
This makes HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects.\cite{Olson2018}
For an allowed electronic transition, the brightness of a vibronic resonance is controlled by overlap of ground and excited state vibrational wavefunctions (Franck-Condon factors).
The $A$ term relies on excited state perturbations to the vibrational wavefunctions; if the excited electronic state minimally changes potential energy surface, so that ground and excited state vibrational eigenfunctions are identical, $\langle 1 | v^\prime \rangle \langle v^\prime | 0 \rangle = 0$ and $A$ vanishes.
%todo: note this is the OPPOSITE of hyper raman 
Some $B$ terms do survive in the same limit, however; for instance, $B_1$ survives for $v^\prime=1$ and $B_2$ survives for $v^\prime=0$.
In the absence of an $A$ term, the $B$ terms will dominate.
The $B_1$ and $B_2$ terms involve different types of vibronic coupling pathways.
In $B_1$, the Herzberg-Teller coupling is associated with the two-photon absorption process, whereas in $B_2$, the Herzberg-Teller coupling is associated with one-photon emission.

The signatures of vibronic coupling in the HDFG spectrum is made clear by using a simple model.
Using a two-well system described by 
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\xi)^2 \right) + \hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ (momentum) and $q$ (position) are the dimensionless conjugate coordinates of the normal mode, the A and B contributions can be evaluated using Franck-Condon and Herzberg-Teller integrals tabulated in terms of the dimensionless normal mode $q$ and offset parameter $\xi=0.5$ (\autoref{fig:doubres_spec}a).\cite{Carlson1988thesis} % DDK: there is probably a better citation than a thesis.  Carlson wasn't the first to derive these...
For simplicity, we restrict consideration to $v^\prime \in \{0,1,2\}$.
Note that $q = \sqrt{\frac{m^2\omega}{\hbar}} Q$.
We assume the electronic state is both one and two photon allowed.

@@

\autoref{fig:doubres_spec}b shows the simulated $\omega_2+\omega_3$ spectrum for this two-well system.  The spectrum selectively probes vibronic pathways from $\ket{g,0}$ to $\ket{g,1}$.
In the resonant region, the $A$ and $B$ terms contain character from all three possible transitions between $\ket{g,0}$, $\ket{e,v'}$ and $\ket{g,1}$. 
% This discussion assumes that mode symmetry allows for coupling between all of the vibronic states considered here.  ddk: there is only one mode!
Significant contributions from $\ket{e,0}$ and $\ket{e,1}$ in the $A$ term spectrum correspond well with expectations given the relative wavefunction overlap in \autoref{fig:doubres_spec}a, tabulated in \autoref{T:contributions}. 
Similarly, the $B_1$ term is dominated by coupling with the $\ket{e,1}$ state and $B_2$ is dominated by coupling with the $\ket{e,0}$ state (\autoref{T:contributions}). 
Note that contributions involving $\ket{e,2}$ are suppressed relative to the $\ket{e,0}$ and $\ket{e,1}$ contributions in the $A$ and $B$ terms.
This is expected, considering the large change in quantum number when involving $\ket{e,2}$ and the offset of the potential wells in \autoref{Hamiltonian}.

>>>>

\documentclass[aip, jcp, reprint, onecolumn, nofootinbib]{revtex4-2}

@@

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares them with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require macroscopic non-centrosymmetry under the electric dipole approximation.\cite{Shen94}
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to four-wave mixing (FWM) cascades.\cite{RN515, RN243, Cho2000_Cascade}
A coherent six-wave mixing analogue of hyper-Raman scattering, coherent anti-Stokes hyper-Raman spectroscopy (CAHRS), has been proposed and explored theoretically by several authors; however, it is also weak and highly susceptible to FWM cascades.\cite{Berger1978, Bjarnason1980, Cho1997, Cho1998}
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave and six-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just a single vibrational mode and electronic state (``evx'' coupling).\cite{RN445, RN335} 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as two-color experiments (with $\omega_2=\omega_3$);\cite{Cho2001} vibrational sum frequency generation (vSFG) or difference frequency generation (vDFG) setups can be trivially reconfigured to perform HSFG or HDFG measurements, respectively, to investigate ev coupling in bulk systems.

When the hyper-Raman excitation in HDFG or HSFG is resonant or near-resonant with electronic states, the brightness of the vibrational feature depends on the vibronic structure of the electronic state.
The electronic spectrum probed in HDFG or HSFG can inform on processes that control ultrafast electronic relaxation in molecular and biological systems in similar ways to resonance Raman.\cite{Cho2001, Bredenbeck2015, Arsenault2021}
Since the selection rules of hyper-Raman scattering differ from Raman scattering, the hyper-Raman excitation spectra give a unique alternative to Raman scattering to understand vibrational spectra, electronic structure and vibronic coupling in molecular systems. \cite{Olson2018}

@@

\section{Density Matrix Propagation}
In the cases discussed in the introduction, the proposed HDFG spectroscopy method is used to probe vibronic coupling.
Coherent spectroscopies are often described using the density-matrix formalism, where a density matrix $\rho$ is propagated using the Liouville-von Neumann equation as
\begin{equation}
	\dot{\rho} = \frac{1}{i \hbar} [H_0 + H(t), \rho] + \dot{\rho}_R,
\end{equation}
where $H_0$ is system Hamiltonian, $H(t)$ is the time-dependent, matter-radiation interaction Hamiltonian and $\rho_R$ is the relaxation density matrix. 
In most treatments of coherent vibrational spectroscopy, it is sufficient to write $\dot{\rho}_R = \Gamma \rho$, i.e. exponential relaxation (Markovian dynamics), where $\Gamma$ is the system's pure dephasing rate. \cite{RN455, RN307}

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
HSFG will not be discussed, but the application of our treatment to HSFG is straightforward and the selection rules are similar.
We note that HDFG emission can be phase-matched in media with normal dispersion, while HSFG cannot,\cite{RN278} although experimental strategies can be implemented to minimize phase-mismatch in HSFG experiments. \cite{Kaufman2024}

It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.\cite{RN287}
While we have only isolated one specific HDFG resonance and time-ordering scheme (\autoref{fig:comparisonwmel}), there are several other pathways which can provide HDFG output. \cite{RN352}
If $\omega_2$ becomes resonant and $\omega_1$ becomes non-resonant, where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the pathways presented in \autoref{fig:sivewmel2} will interfere, which can amplify or obscure coherent output.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66 in]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$, described in \autoref{Albrecht}.
		In all cases, $\omega_2 > \omega_1$.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ is detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ interacts first and is resonant, and $\omega_2 + \omega_3$ is nonresonant, yields two contributed pathways: \autoref{fig:sivewmel2} (a) and (b).
These pathways possess identical phase and transition moments, which yields total constructive interference, assuming similar non-resonant dispersion from each pathway.
Conversely, in the case where only $\omega_2$ is vibrationally resonant, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned. \cite{RN287, McDonnell2024}
Note that in the case where $\omega_1$ interacts first but $\omega_2$ is resonant, the nonlinear output generated nearly vanishes as the first resonant pulse arrives after a non-resonant pulse.

@@

It is useful to expose relationships between nonlinear output, transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit. \cite{Simpson2004, Moad2005}
Under the electric dipole approximation, the $I^\text{th}$ cartesian component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation) \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(\omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the $IJKL$ element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}\label{eq:nfgamma}
	\chi^{(3)}_{IJKL} = NF \langle \gamma_{ijkl} \rangle.
\end{equation}
Here $N$ is a number density and $\gamma_{ijkl}$ is the third-order polarizability (i.e., second hyperpolarizability).
The brackets indicate an orientational average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.
$F=\prod_j  \frac{1}{3} \left({\left(n(\omega_j)\right)^2 + 1} \right)$ is the Lorentz local field factor, where $n$ is the (frequency dependent) refractive index. \cite{Bedeaux1973}

@@

\subsection{Vibronic Selection Rules}\label{Albrecht}
While the Placzek treatment provides the general source of HDFG output, it does not predict the behavior of $\beta_{ijk}$ as $\omega_2 + \omega_3$ is changed.
We now explicitly consider electronic resonances through the $A,B,C$ decomposition of $\beta$ introduced by Chung and Ziegler, analogous to those found in Albrecht's treatment of Raman spectroscopy.\cite{Albrecht1961, Ziegler1988} 
To employ this formalism, we write the states in a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a(Q)) \otimes \ket{b}$ for electronic states $\{|a(\textbf{Q}))\}$ and vibrational states $\{\ket{b}\}$ (i.e., adiabatic approximation), where \textbf{Q}, the system normal modes, is the adiabatic parameter. \cite{BornOppenheimer, Tang1970}
We will henceforth suppress the dependence of the electronic states on \textbf{Q} for simplicity.

For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments; $\vec{\mu}$ is reserved for transitions on the ground electronic state. \cite{Tang1970}
Following the approach which gave \autoref{sivegamma} and using the time ordering and state labeling of \autoref{fig:sivewmel2}a, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{4D \hbar^3} \sum_{m,n,e,v'} \frac{
		R_{i}^{gv, ev'} 
		R_{j}^{ev',mn} 
		R_{k}^{mn,g0} 
		R_{l}^{g0,gv} 
	}{	\Delta_{ev', gv}
		\Delta_{mn, gv}
		\Delta_{g0,gv}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.
It is common in the hyper-Raman literature to write $\Delta_{ev', gv} \Delta_{mn, gv} = \Delta_{ev', gv}$.
Using our definition of vibronic states, we can write, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$, where $\vec{M}^{ab} = (a|\vec{R}|b)$.\cite{Ziegler1974}
Expanding $\vec{M}^{ij}$ to $\order{Q}$ about the equilibrium point of the ground state potential surface as
$\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z} Q_z$
yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Warshel1977, Ziegler1988} so that
\begin{equation}\label{gammaabc}
		\gamma_{ijkl} \sim \left(A_{ijk} + B_{ijk} + C_{ijk}\right) \frac{\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}
\end{equation}
The $A$ term contains the static ($\order{Q^0}$) transitions (i.e., Condon approximation), the $B$ term depends upon $\order{Q}$ transitions (Herzberg-Teller contributions), and the $C$ term depends on $\order{Q_i Q_j}$ transitions. 
The $C$ term is suppressed in the following discussion as it depends on one and two photon forbidden transitions, making its contribution to $\gamma_{ijkl}$ roughly two orders of magnitude less than $A_{ijk}$. \cite{Ziegler1988, Neddersen1989, Bonang1992}
Note that some reports obtain $A, B$ coefficients through a Herzberg-Teller expansion of the electronic states to expose vibronic couplings via $\partial H / \partial Q$, where $H$ is the electronic Hamiltonian.\cite{HerzbergTeller1933, Petrov1985, Neddersen1989, Baranov1990}
This approach is not used here as the expansion of $\vec{M}^{ij}$ in normal mode coordinates provides sufficient physical insight into hyper-Raman selection rules.

The $A$ and $B$ coefficients, where $B = B_1 + B_2$, following closure over $\ket{n}$,\cite{Milojevich2013} are written as:
\begin{widetext}
\begin{subequations}\label{ABterms}
\begin{equation}
	\begin{split}
		A_{ijk} = \frac{1}{\hbar^2}\sum_{m,e,v'} M^{ge}_{0,i} 
		M^{em}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | v' \rangle
		 \langle v' | 0 \rangle 
		 \frac{1}{\Delta_{ev', gv}}
		 \\
	\end{split}
\end{equation}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \frac{1}{\hbar^2} \sum_{m,e,v',z} M^{ge}_{0,i} \left(
				\frac{\partial M^{em}_{j}}{\partial Q_z} M^{mg}_{0,k}  
				+M^{em}_{0,j} \frac{\partial M^{mg}_{k}}{\partial Q_z}
			\right)
			\langle v | v' \rangle \mel{v'}{Q_z}{0} \frac{1}{\Delta_{ev', gv}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \frac{1}{\hbar^2} \sum_{m,e,v',z} \frac{\partial M^{ge}_{i}}{\partial Q_z} M^{em}_{0,j} 
			M^{mg}_{0,k} \mel{v}{Q_z}{v'} 
			\langle v' | 0 \rangle 
			\frac{1}{\Delta_{ev', gv}}
	\end{split}
	\end{equation}
\end{subequations}
\end{widetext}
where $\langle a | b \rangle$ and $\mel{a}{Q}{b}$ are Franck-Condon factors and Herzberg-Teller-type integrals, respectively. 
Duschinsky rotation effects are ignored in this discussion. \cite{Duschinsky1937}
At this point, the expressions are valid for both electronically resonant and non-resonant cases.

To understand the vibronic structure of electronic resonance, we restrict consideration to electronic transitions between $|g)$ and $|e)$.  
Note that HDFG experiments selectively excite ground state vibrational modes through $\omega_1$; we will assume the IR beam has excited the fundamental of an IR-active mode so that hyper-Raman scattering is between this mode and the ground state. 
By working only in terms of this single normal mode (i.e., removing the sum over $z$ and taking $v=1$), 
the $A$ and $B$ terms are rewritten as 
	\begin{subequations}\label{ABterms_DR}
		\begin{equation}
			\begin{split}
				A_{ijk} = \frac{1}{\hbar}\sum_{v'} M^{ge}_{0,i} 
				\Lambda^{eg}_{0,jk}
				\langle 1 | v' \rangle
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',gv}}
				\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{1_{ijk}} &= \frac{1}{\hbar} \sum_{v'} M^{ge}_{0,i}
				\frac{\partial \Lambda^{eg}_{jk}}{\partial Q} \langle 1 | v' \rangle \mel{v'}{Q}{0} 
				\frac{1}{\Delta_{ev', gv}}\\
			\end{split}
		\end{equation}
		\begin{equation}
			\begin{split}
				B_{2_{ijk}} = \frac{1}{\hbar} \sum_{v'} \frac{\partial M^{ge}_{i}}{\partial Q} 
				\Lambda^{eg}_{0,jk} 
				\mel{1}{Q}{v'} 
				\langle v' | 0 \rangle 
				\frac{1}{\Delta_{ev',gv}}
			\end{split}
		\end{equation}
	\end{subequations}
where $\Lambda^{ab}_{ij} = 1/\hbar \sum_m M_i^{am}M_j^{mb} $ is the electronic two photon transition moment between $a$ and $b$.\cite{McClain1977}

@@

The signatures of vibronic coupling in the HDFG spectrum are made clear by using a simple model.
Using a two-well system described by 
\begin{subequations}\label{Hamiltonian}
	\begin{equation}
		H_g = \frac{\hbar \omega }{2} \left(p^2 + q^2 \right)
	\end{equation}
	and
	\begin{equation}
		H_e = \frac{\hbar \omega }{2} \left(p^2 +  (q-\xi)^2 \right) + \hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ (momentum) and $q$ (position) are the dimensionless conjugate coordinates of the normal mode, the A and B contributions can be evaluated using Franck-Condon and Herzberg-Teller integrals tabulated in terms of the dimensionless normal mode $q$ and offset parameter $\xi$ (\autoref{fig:doubres_spec}a).\cite{Carlson1988thesis} 
For simplicity, we restrict consideration to $v^\prime \in \{0,1,2\}$ and take $\xi = 0.5$.
Note that $q = \sqrt{\frac{m^2\omega}{\hbar}} Q$.
We assume the electronic state is both one and two photon allowed.

\autoref{ABterms_DR} shows important selection rules for the vibronic transitions.
The $A$ term is allowed whenever an electronic transition has a one- and two- photon transition moment.
For centrosymmetric molecules, an electronic transition cannot be both one- and two- photon allowed by parity, so the $A$ term will vanish.
This makes HDFG a unique tool for studying the electronic structure of centrosymmetric species in isotropic media, as it is sensitive to non-Condon effects.\cite{Olson2018}
For an allowed electronic transition, the brightness of a vibronic resonance is controlled by overlap of ground and excited state vibrational wavefunctions (Franck-Condon factors).
The $A$ term relies on excited state perturbations to the vibrational wavefunctions; if the excited electronic state minimally changes potential energy surface, so that ground and excited state vibrational eigenfunctions are identical, $\langle 1 | v^\prime \rangle \langle v^\prime | 0 \rangle = 0$ and $A$ vanishes.
This follows from $\lim_{\xi \rightarrow 0} \langle v' | 0 \rangle = 0$, $v^\prime \neq 0$, for wells with identical curvature.
Some $B$ terms do survive in the same limit, however; for instance, $B_1$ survives for $v^\prime=1$ and $B_2$ survives for $v^\prime=0$.
In the absence of an $A$ term, the $B$ terms will dominate.
The $B_1$ and $B_2$ terms involve different types of vibronic coupling pathways.
In $B_1$, the Herzberg-Teller coupling is associated with the two-photon absorption process, whereas in $B_2$, the Herzberg-Teller coupling is associated with one-photon emission.
